---
title: "Interrupted Time Series Analysis"
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 3
    embed-resources: true
---

# Overview

This analysis examines two major interventions that reshaped the sports betting landscape: the COVID-19 pandemic (March 2020) and New York's online sports betting legalization (January 2022). Using Google Trends data spanning 2020-2024, we employ interrupted time series regression to quantify both immediate shocks and sustained trend changes. Google Trends provides a direct cultural proxy for public engagement, capturing how these events transformed sports betting from a niche activity during COVID's sports shutdown to mainstream entertainment following New York's market entry. The ITS model 

> $Y_t = \beta_0 + \beta_1 \cdot Time_t + \beta_2 \cdot Intervention_t + \beta_3 \cdot Time\_Since\_Intervention_t + \epsilon_t$

isolates intervention effects from underlying trends, measuring level changes ($\beta_2$) and slope changes ($\beta_3$) in search interest. 

---

# Data & Setup

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(forecast)
library(gridExtra)
library(kableExtra)

theme_set(theme_minimal(base_size = 12))

# Load Google Trends data
trends <- read_csv("data/google_trends/sports_betting_trends.csv", show_col_types = FALSE) %>%
    mutate(date = as.Date(date))

# Create separate dataframes and add ITS variables
intervention_date <- as.Date("2022-01-08")
covid_intervention <- as.Date("2020-03-11")

add_its_variables <- function(df, int_date) {
    df %>%
        mutate(
            Post_Intervention = ifelse(date >= int_date, 1, 0),
            Time = as.numeric(date - min(date)),
            Time_Since_Intervention = ifelse(Post_Intervention == 1,
                as.numeric(date - int_date), 0
            )
        )
}

sports_betting <- trends %>%
    filter(keyword == "sports betting") %>%
    add_its_variables(intervention_date)
draftkings <- trends %>%
    filter(keyword == "draftkings") %>%
    add_its_variables(intervention_date)
online_betting <- trends %>%
    filter(keyword == "online betting") %>%
    add_its_variables(intervention_date)
nba_betting <- trends %>%
    filter(keyword == "nba betting") %>%
    add_its_variables(intervention_date)
```

---

::: {.panel-tabset}

## NY Legalization Analysis

### Background & Data

**Intervention**: New York online sports betting launch (January 8, 2022) - the largest U.S. sports betting market by revenue, representing ~20% of the national market. 

**Research Question**: Did NY legalization produce a significant change in public interest in sports betting? 

> We analyze Google search trends as a direct measure of cultural impact and consumer awareness, examining four search terms that capture different facets of betting interest from general ("sports betting") to operator-specific ("draftkings") to sport-specific ("nba betting").

### Visualization & Data Preparation

```{r ny-viz-data, warning=FALSE, message=FALSE, fig.width=14, fig.height=10}
p1 <- ggplot(sports_betting, aes(x = date, y = hits)) +
    geom_line(color = "#1e88e5", linewidth = 1) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.2) +
    annotate("text",
        x = intervention_date, y = max(sports_betting$hits) * 0.95,
        label = "NY Launch\nJan 8, 2022", color = "red", size = 4, fontface = "bold"
    ) +
    labs(
        title = "Google Searches: 'sports betting'",
        subtitle = "NY sports betting legalization intervention point marked in red",
        x = "Date", y = "Search Interest (0-100)"
    ) +
    theme(plot.title = element_text(face = "bold"))

p2 <- ggplot(draftkings, aes(x = date, y = hits)) +
    geom_line(color = "#ff6f00", linewidth = 1) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.2) +
    annotate("text",
        x = intervention_date, y = max(draftkings$hits) * 0.95,
        label = "NY Launch\nJan 8, 2022", color = "red", size = 4, fontface = "bold"
    ) +
    labs(title = "Google Searches: 'draftkings'", x = "Date", y = "Search Interest (0-100)") +
    theme(plot.title = element_text(face = "bold"))

p3 <- ggplot(online_betting, aes(x = date, y = hits)) +
    geom_line(color = "#43a047", linewidth = 1) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.2) +
    annotate("text",
        x = intervention_date, y = max(online_betting$hits) * 0.95,
        label = "NY Launch\nJan 8, 2022", color = "red", size = 4, fontface = "bold"
    ) +
    labs(title = "Google Searches: 'online betting'", x = "Date", y = "Search Interest (0-100)") +
    theme(plot.title = element_text(face = "bold"))

p4 <- ggplot(nba_betting, aes(x = date, y = hits)) +
    geom_line(color = "#8e24aa", linewidth = 1) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.2) +
    annotate("text",
        x = intervention_date, y = max(nba_betting$hits) * 0.95,
        label = "NY Launch\nJan 8, 2022", color = "red", size = 4, fontface = "bold"
    ) +
    labs(title = "Google Searches: 'nba betting'", x = "Date", y = "Search Interest (0-100)") +
    theme(plot.title = element_text(face = "bold"))

gridExtra::grid.arrange(p1, p2, p3, p4, ncol = 2)

# Data preparation table
prep_table <- sports_betting %>%
    filter(date >= as.Date("2021-12-01") & date <= as.Date("2022-02-28")) %>%
    select(date, hits, Time, Post_Intervention, Time_Since_Intervention) %>%
    mutate(Period = ifelse(Post_Intervention == 0, "Pre-NY", "Post-NY")) %>%
    select(date, Period, Y = hits, X_t = Time, Z_t = Post_Intervention, P_t = Time_Since_Intervention)

kable(prep_table,
    format = "html", digits = 1,
    caption = "ITS Variables for NY Legalization Analysis (Sample Period)",
    col.names = c("Date", "Period", "Y (Search Interest)", "X_t (Time)", "Z_t (NY Indicator)", "P_t (Time Since NY)")
) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
    row_spec(which(prep_table$Z_t == 1), background = "#e3f2fd")
```

### ITS Model & Results

**Model**: $Y_t = \beta_0 + \beta_1 X_t + \beta_2 Z_t + \beta_3 P_t + \varepsilon_t$
, where $\beta_1$ captures pre-intervention trend, $\beta_2$ measures immediate level change, and $\beta_3$ quantifies trend change after intervention.

```{r ny-its-models, warning=FALSE, message=FALSE, fig.width=12, fig.height=6}
# Fit models for all search terms
model_sb <- lm(hits ~ Time + Post_Intervention + Time_Since_Intervention, data = sports_betting)
model_dk <- lm(hits ~ Time + Post_Intervention + Time_Since_Intervention, data = draftkings)
model_ob <- lm(hits ~ Time + Post_Intervention + Time_Since_Intervention, data = online_betting)
model_nb <- lm(hits ~ Time + Post_Intervention + Time_Since_Intervention, data = nba_betting)

# Extract coefficients
extract_results <- function(model, name) {
    coef <- summary(model)$coefficients
    data.frame(
        Term = name,
        Parameter = c("Intercept", "Pre-Trend (β₁)", "Level Change (β₂)", "Trend Change (β₃)"),
        Coefficient = coef[, 1],
        Std_Error = coef[, 2],
        p_value = coef[, 4]
    )
}

results_sb <- extract_results(model_sb, "Sports Betting")
results_dk <- extract_results(model_dk, "DraftKings")
results_ob <- extract_results(model_ob, "Online Betting")
results_nb <- extract_results(model_nb, "NBA Betting")

all_results <- rbind(results_sb, results_dk, results_ob, results_nb)

kable(all_results,
    format = "html", digits = 4,
    caption = "NY Legalization ITS Results: All Search Terms",
    col.names = c("Term", "Parameter", "Coefficient", "Std. Error", "p-value")
) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
    row_spec(which(all_results$p_value < 0.05), background = "#d4edda") %>%
    pack_rows("Sports Betting", 1, 4) %>%
    pack_rows("DraftKings", 5, 8) %>%
    pack_rows("Online Betting", 9, 12) %>%
    pack_rows("NBA Betting", 13, 16)

# Fitted values plot
sports_betting$Fitted <- fitted(model_sb)

ggplot(sports_betting, aes(x = date)) +
    geom_line(aes(y = hits), color = "gray60", alpha = 0.6, linewidth = 0.5) +
    geom_line(aes(y = Fitted), color = "#1e88e5", linewidth = 1.2) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.2) +
    annotate("text",
        x = intervention_date, y = max(sports_betting$hits) * 0.95,
        label = "NY Launch", color = "red", size = 5, fontface = "bold"
    ) +
    labs(
        title = "'Sports Betting' Searches: Actual vs. Fitted ITS Trend Lines",
        subtitle = "Gray = Actual | Blue = Fitted (shows pre- and post-intervention trends)",
        x = "Date", y = "Search Interest (0-100)"
    ) +
    theme(plot.title = element_text(face = "bold", size = 15))
```

### Counterfactual & Interpretation

```{r ny-counterfactual, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
# Generate predictions and counterfactuals
sports_betting <- sports_betting %>%
    mutate(
        Predicted = predict(model_sb),
        Counterfactual = predict(model_sb, newdata = sports_betting %>% mutate(Post_Intervention = 0, Time_Since_Intervention = 0)),
        Effect = Predicted - Counterfactual
    )

# Counterfactual visualization
ggplot(sports_betting, aes(x = date)) +
    geom_line(aes(y = Counterfactual, linetype = "Counterfactual (No NY)"), color = "#1e88e5", linewidth = 1.2) +
    geom_line(aes(y = Predicted, linetype = "Predicted (With NY)"), color = "#d32f2f", linewidth = 1.2) +
    geom_point(aes(y = hits, shape = "Actual Data"), color = "gray30", size = 1, alpha = 0.5) +
    geom_ribbon(
        data = sports_betting %>% filter(Post_Intervention == 1),
        aes(ymin = Predicted, ymax = Counterfactual), fill = "red", alpha = 0.15
    ) +
    geom_vline(xintercept = intervention_date, color = "red", linetype = "dashed", linewidth = 1.5) +
    annotate("text",
        x = intervention_date, y = max(sports_betting$hits) * 0.98,
        label = "NY Launch", color = "red", size = 5, fontface = "bold", hjust = -0.1
    ) +
    scale_linetype_manual(
        name = "Trend Lines",
        values = c("Counterfactual (No NY)" = "dashed", "Predicted (With NY)" = "solid")
    ) +
    scale_shape_manual(name = "Data", values = c("Actual Data" = 16)) +
    labs(
        title = "Predicted vs. Counterfactual: Impact of NY Sports Betting Legalization",
        subtitle = "Red shaded area shows intervention effect (difference between what happened and what would have happened)",
        x = "Date", y = "Search Interest (0-100)",
        caption = "Blue dashed = Counterfactual (pre-NY trend continued) | Red solid = Predicted (with NY intervention)"
    ) +
    theme(plot.title = element_text(face = "bold", size = 16), legend.position = "bottom")

# Coefficient interpretation
coefs <- coef(model_sb)
coef_summary <- summary(model_sb)$coefficients

# Delayed effect
post_ny <- sports_betting %>% filter(Post_Intervention == 1)
avg_effect <- mean(post_ny$Effect)
current_effect <- tail(sports_betting$Effect, 1)
counterfactual_current <- tail(sports_betting$Counterfactual, 1)
actual_current <- tail(sports_betting$hits, 1)
```

**Coefficient Interpretation**

- **β₁ (Pre-Trend)** = `r sprintf("%.4f", coefs[2])` (p = `r sprintf("%.4f", coef_summary[2, 4])`): Search interest `r ifelse(coefs[2] > 0, "increased", "decreased")` by `r sprintf("%.4f", abs(coefs[2]))` points per week before NY launch `r ifelse(coef_summary[2, 4] < 0.05, "(significant)", "(not significant)")`

- **β₂ (Level Change)** = `r sprintf("%.4f", coefs[3])` (p = `r sprintf("%.4f", coef_summary[3, 4])`): `r ifelse(coef_summary[3, 4] < 0.05, "Significant", "No significant")` immediate `r ifelse(coefs[3] > 0, "jump", "drop")` of `r sprintf("%.2f", abs(coefs[3]))` points when NY launched `r ifelse(coef_summary[3, 4] < 0.05, "(intervention had immediate effect)", "(no immediate shock)")`

- **β₃ (Trend Change)** = `r sprintf("%.4f", coefs[4])` (p = `r sprintf("%.4f", coef_summary[4, 4])`): Post-NY trend = `r sprintf("%.4f", coefs[2])` (pre) + `r sprintf("%.4f", coefs[4])` (change) = `r sprintf("%.4f", coefs[2] + coefs[4])` per week
  `r ifelse(coefs[4] > 0, "Acceleration", "Deceleration")` in growth rate after NY launch `r ifelse(coef_summary[4, 4] < 0.05, "(significant sustained effect)", "(not significant)")`

- **Average Effect (Post-NY)**: `r sprintf("%.2f", avg_effect)` points
- **Current Effect**: `r sprintf("%.2f", current_effect)` points

**Counterfactual**: Without NY legalization, current search interest would be `r sprintf("%.1f", counterfactual_current)` instead of `r sprintf("%.1f", actual_current)`

The β₂ coefficient captures the immediate effect, indicating whether search interest jumped or declined right after New York’s January 8, 2022 legalization, with a significant negative value signaling an immediate cultural response. The counterfactual blue dashed line shows what search interest would have been without the intervention, and the gap between it and the actual red line quantifies the overall impact at any point. The β₃ coefficient reflects the sustained effect by measuring how the growth trajectory changed post-legalization, where negative values indicate slowing interest as the market stabilized. Together with the delayed effect, intervention difference and the ongoing separation between actual and counterfactual, these patterns reveal how the impact evolved and persisted well beyond the initial launch period.

## COVID-19 Analysis

### Background & Variable Selection

**Intervention**: WHO declared COVID-19 pandemic on March 11, 2020, causing immediate suspension of all major sports leagues (NBA, NHL, MLB). 

**Variable**: "Sports betting" Google searches - an ITS candidate because the intervention had a clear, direct causal pathway: no live sports = no sports betting.

### Visualization & Data Preparation

```{r covid-viz-data, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
# Prepare COVID ITS data
sports_covid_its <- sports_betting %>%
    mutate(
        X_t = row_number(),
        Z_t = ifelse(date >= covid_intervention, 1, 0),
        P_t = ifelse(date >= covid_intervention, as.numeric(date - covid_intervention) / 7, 0),
        Y = hits
    )

# Visualization
sports_covid <- sports_covid_its %>%
    filter(date >= as.Date("2019-12-01") & date <= as.Date("2021-12-31"))

p_covid_impact <- ggplot(sports_covid, aes(x = date, y = Y)) +
    geom_line(color = "#1e88e5", linewidth = 1.2) +
    geom_point(color = "#1e88e5", size = 1.5, alpha = 0.6) +
    geom_vline(xintercept = covid_intervention, color = "red", linetype = "dashed", linewidth = 1.5) +
    annotate("text",
        x = covid_intervention, y = max(sports_covid$Y) * 0.95,
        label = "COVID-19 Pandemic\nMarch 11, 2020",
        color = "red", size = 5, fontface = "bold", hjust = -0.1
    ) +
    annotate("rect",
        xmin = covid_intervention, xmax = as.Date("2020-07-01"),
        ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red"
    ) +
    # annotate("text",
    #     x = as.Date("2020-05-01"), y = max(sports_covid$Y) * 0.1,
    #     label = "Sports Leagues Shut Down",
    #     color = "darkred", size = 4, fontface = "italic"
    # ) +
    labs(
        title = "Google Searches for 'Sports Betting' Around COVID-19 Pandemic",
        subtitle = "Dramatic decline when sports leagues shut down in March 2020",
        x = "Date", y = "Search Interest (0-100)",
        caption = "Red shaded area indicates major sports leagues suspended"
    ) +
    theme(plot.title = element_text(face = "bold", size = 16))

# Display plot
print(p_covid_impact)

# Save for conclusion
ggsave("images/covid_betting_impact.png", plot = p_covid_impact, width = 14, height = 8, dpi = 300, bg = "white")

# Data preparation table
covid_table <- sports_covid_its %>%
    filter(date >= as.Date("2020-02-01") & date <= as.Date("2020-05-31")) %>%
    select(date, Y, X_t, Z_t, P_t) %>%
    mutate(Period = ifelse(Z_t == 0, "Pre-COVID", "Post-COVID"), P_t = round(P_t, 1)) %>%
    select(date, Period, Y, X_t, Z_t, P_t)

kable(covid_table,
    format = "html", digits = 1,
    caption = "ITS Variables for COVID-19 Analysis (Sample Period)",
    col.names = c("Date", "Period", "Y (Search Interest)", "X_t (Time)", "Z_t (COVID Indicator)", "P_t (Weeks Since COVID)")
) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    row_spec(0, bold = TRUE, background = "#2c3e50", color = "white") %>%
    row_spec(which(covid_table$Z_t == 1), background = "#ffebee") %>%
    column_spec(5, bold = TRUE, color = ifelse(covid_table$Z_t == 1, "red", "black"))
```

### ITS Model & Results

```{r covid-its-model, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
# Fit COVID ITS model
covid_model <- lm(Y ~ X_t + Z_t + P_t, data = sports_covid_its)

covid_coef <- summary(covid_model)$coefficients

covid_results <- data.frame(
    Parameter = c(
        "β₀: Intercept (Baseline)", "β₁: Pre-COVID Trend (X_t)",
        "β₂: Immediate Level Change (Z_t)", "β₃: Trend Change After COVID (P_t)"
    ),
    Coefficient = covid_coef[, 1],
    Std_Error = covid_coef[, 2],
    t_value = covid_coef[, 3],
    p_value = covid_coef[, 4],
    Significance = ifelse(covid_coef[, 4] < 0.001, "***",
        ifelse(covid_coef[, 4] < 0.01, "**",
            ifelse(covid_coef[, 4] < 0.05, "*", "")
        )
    )
)

kable(covid_results,
    format = "html", digits = 4,
    caption = "COVID-19 ITS Model Results: Sports Betting Search Interest",
    col.names = c("Parameter", "Coefficient", "Std. Error", "t-value", "p-value", "Sig.")
) %>%
    kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed")) %>%
    row_spec(0, bold = TRUE, background = "#c62828", color = "white") %>%
    row_spec(which(covid_results$p_value < 0.05), background = "#ffcdd2")

# Fitted values visualization
sports_covid_its$Fitted <- fitted(covid_model)

ggplot(sports_covid_its, aes(x = date)) +
    geom_line(aes(y = Y), color = "gray50", linewidth = 0.8, alpha = 0.7) +
    geom_point(aes(y = Y), color = "gray50", size = 1, alpha = 0.5) +
    geom_line(
        data = sports_covid_its %>% filter(Z_t == 0), aes(y = Fitted),
        color = "#1e88e5", linewidth = 2
    ) +
    geom_line(
        data = sports_covid_its %>% filter(Z_t == 1), aes(y = Fitted),
        color = "#d32f2f", linewidth = 2
    ) +
    geom_vline(xintercept = covid_intervention, color = "red", linetype = "dashed", linewidth = 1.5) +
    annotate("text",
        x = covid_intervention, y = max(sports_covid_its$Y) * 0.95,
        label = "COVID-19\nMarch 11, 2020", color = "red", size = 5, fontface = "bold", hjust = -0.1
    ) +
    labs(
        title = "ITS Model: Actual Data vs. Fitted Trend Lines",
        subtitle = "Gray = Actual | Blue = Pre-COVID trend | Red = Post-COVID trend",
        x = "Date", y = "Search Interest (0-100)",
        caption = "Model: Y = β₀ + β₁(Time) + β₂(COVID Indicator) + β₃(Weeks Since COVID)"
    ) +
    theme(plot.title = element_text(face = "bold", size = 16))
```

### Counterfactual & Interpretation

```{r covid-counterfactual, warning=FALSE, message=FALSE, fig.width=14, fig.height=9}
# Generate predictions and counterfactuals
sports_covid_its <- sports_covid_its %>%
    mutate(
        Predicted = predict(covid_model),
        Counterfactual = predict(covid_model, newdata = sports_covid_its %>% mutate(Z_t = 0, P_t = 0)),
        Effect = Predicted - Counterfactual
    )

# Counterfactual visualization
ggplot(sports_covid_its, aes(x = date)) +
    geom_line(aes(y = Counterfactual, linetype = "Counterfactual (No COVID)"), color = "#1e88e5", linewidth = 1.2) +
    geom_line(aes(y = Predicted, linetype = "Predicted (With COVID)"), color = "#d32f2f", linewidth = 1.2) +
    geom_point(aes(y = Y, shape = "Actual Data"), color = "gray30", size = 1.5, alpha = 0.5) +
    geom_ribbon(
        data = sports_covid_its %>% filter(Z_t == 1),
        aes(ymin = Predicted, ymax = Counterfactual), fill = "red", alpha = 0.15
    ) +
    geom_vline(xintercept = covid_intervention, color = "red", linetype = "dashed", linewidth = 1.5) +
    annotate("text",
        x = covid_intervention, y = max(sports_covid_its$Y) * 0.98,
        label = "COVID-19\nIntervention", color = "red", size = 5, fontface = "bold", hjust = -0.1
    ) +
    annotate("text",
        x = as.Date("2021-06-01"), y = 40,
        label = "Intervention Effect",
        color = "#c62828", size = 4.5, fontface = "bold"
    ) +
    scale_linetype_manual(
        name = "Trend Lines",
        values = c("Counterfactual (No COVID)" = "dashed", "Predicted (With COVID)" = "solid")
    ) +
    scale_shape_manual(name = "Data", values = c("Actual Data" = 16)) +
    labs(
        title = "Predicted vs. Counterfactual: Impact of COVID-19 on Sports Betting Interest",
        subtitle = "Red shaded area shows intervention effect (difference between actual and counterfactual)",
        x = "Date", y = "Search Interest (0-100)",
        caption = "Blue dashed = Counterfactual (pre-COVID trend) | Red solid = Predicted (with COVID)"
    ) +
    theme(plot.title = element_text(face = "bold", size = 16), legend.position = "bottom")

# Coefficient interpretation
coefs_c <- coef(covid_model)
coef_summary_c <- summary(covid_model)$coefficients

# Effect analysis
post_covid_data <- sports_covid_its %>% filter(Z_t == 1)
avg_effect_c <- mean(post_covid_data$Effect)
min_effect_c <- min(post_covid_data$Effect)
current_effect_c <- tail(sports_covid_its$Effect, 1)
counterfactual_current_c <- tail(sports_covid_its$Counterfactual, 1)
actual_current_c <- tail(sports_covid_its$Y, 1)
```

**Coefficient Interpretation**

- **β₁ (Pre-COVID Trend)** = `r sprintf("%.4f", coefs_c[2])` (p = `r sprintf("%.4f", coef_summary_c[2, 4])`): Search interest `r ifelse(coefs_c[2] > 0, "increased", "decreased")` by `r sprintf("%.4f", abs(coefs_c[2]))` points per week before pandemic

- **β₂ (Immediate Shock)** = `r sprintf("%.4f", coefs_c[3])` (p = `r sprintf("%.4f", coef_summary_c[3, 4])`): Search interest DROPPED by `r sprintf("%.2f", abs(coefs_c[3]))` points when sports stopped. This represents the immediate collapse in betting interest when all sports leagues shut down

- **β₃ (Recovery Rate)** = `r sprintf("%.4f", coefs_c[4])` (p = `r sprintf("%.4f", coef_summary_c[4, 4])`): Post-COVID trend = `r sprintf("%.4f", coefs_c[2])` (pre) + `r sprintf("%.4f", coefs_c[4])` (change) = `r sprintf("%.4f", coefs_c[2] + coefs_c[4])` per week
  Trend became `r ifelse(coefs_c[2] + coefs_c[4] > coefs_c[2], "MORE POSITIVE", "MORE NEGATIVE")`, indicating `r ifelse(coefs_c[4] > 0, "RECOVERY", "continued decline")` as sports returned

- **Average Effect (Post-COVID)**: `r sprintf("%.2f", avg_effect_c)` points below counterfactual
- **Worst Point**: `r sprintf("%.2f", min_effect_c)` points below counterfactual (peak disruption)
- **Current Effect**: `r sprintf("%.2f", current_effect_c)` points

**Counterfactual**: Without COVID, current search interest would be `r sprintf("%.1f", counterfactual_current_c)` instead of `r sprintf("%.1f", actual_current_c)`

```{r covid-effect-plot, warning=FALSE, message=FALSE, fig.width=14, fig.height=8}
ggplot(post_covid_data, aes(x = date, y = Effect)) +
    geom_line(color = "#d32f2f", linewidth = 1.2) +
    geom_point(color = "#d32f2f", size = 2) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray30", linewidth = 1) +
    geom_smooth(method = "loess", se = TRUE, color = "#1e88e5", fill = "#1e88e5", alpha = 0.2) +
    labs(
        title = "Intervention Effect Over Time: COVID-19 Impact Trajectory",
        subtitle = "Negative values = Interest below counterfactual | Positive = Above counterfactual",
        x = "Date", y = "Intervention Effect (Predicted - Counterfactual)",
        caption = "Blue trend line shows recovery trajectory over time"
    ) +
    theme(plot.title = element_text(face = "bold", size = 15))
```

β₂ captures the immediate effect, showing that search interest dropped roughly %.2 points when sports shut down, a catastrophic collapse that erased the core driver of betting demand, while the counterfactual trajectory shows what interest would have been without COVID, with the gap representing “lost interest,” averaging %.1 points below the expected level. β₃ reflects the sustained effect, revealing how the trend shifted as sports gradually returned with the NBA bubble and NFL season, producing a delayed pattern of sharp collapse followed by slow recovery. The current distance between the actual and counterfactual lines indicates whether the industry has fully realigned with its pre-pandemic path or if meaningful deficits persist.

:::

---
